---
title: "Clean-up Feature Databases"
output: html_document
---

```{r setup, message = FALSE}
# set default global knitr options
knitr::opts_chunk$set(echo = TRUE, results = 'hold', fig.show = 'hold')
library(readr)
library(dplyr)
library(reticulate)
# use_condaenv('rpy3_tmp')
Sys.setenv('RETICULATE_PYTHON'='/home/users/bentyeh/miniconda3/envs/rpy3_tmp/bin/python')
options(stringsAsFactors=FALSE, warn=1)

if (.Platform$OS.type == 'unix') {
    projectDir = normalizePath('~/projects/disprot')
} else {
    projectDir = normalizePath('~/Projects/disprot')
}

files = reticulate::import_from_path(module='files', path=file.path(projectDir, 'scripts'))
paths = files$getPaths()
source(file.path(projectDir, 'scripts', 'disorderedSeqs.R'))
source(file.path(projectDir, 'scripts', 'featurize.R'))
```

Read in original features database

```{r, message = FALSE}
featuresDB_orig = list()
col_types = cols(.default = col_double(), rowname = col_character())
seqtypes = c('full', 'disorder')
lambdas = c(0, 25, 50, 75, 100)
for (seqtype in seqtypes) {
  for (lambda in lambdas) {
    db_name = paste0(seqtype, lambda)
    print(db_name)
    featuresDB_orig[[paste0(seqtype, lambda)]] = readr::read_tsv(
      file = paths[[paste('features', db_name, 'human', 'orig', sep='_')]],
      col_types = col_types)
  }
}
```

Drop rows (proteins) with NA-valued features

```{r}
featuresDB = list()
for (seqtype in seqtypes) {
  for (lambda in lambdas) {
    db_name = paste0(seqtype, lambda)
    featuresDB[[db_name]] = featuresDB_orig[[db_name]] %>% tidyr::drop_na()
    #featuresDB[[db_name]] %>%
    #  readr::write_tsv(path = paths[[paste('features', db_name, 'human', sep='_')]])
  }
}
```

See which sequence identifiers were differentially dropped across features databases

```{r}
for (seqtype in seqtypes) {
  for (lambda in setdiff(lambdas, 0)) {
    db_name = paste0(seqtype, lambda)
    print('--------------------')
    print(db_name)
    db_rownames = rownames(featuresDB[[db_name]])
    db0_rownames = rownames(featuresDB[[paste0(seqtype, 0)]])
    # print('The following sequence identifiers are only available in the length-independent database:')
    # print(setdiff(db0_rownames, db_rownames))
    print('The following sequence identifiers are only available in the length-dependent database:')
    print(setdiff(db_rownames, db0_rownames))
    print('--------------------')
  }
}
```

Create dataframe of locations (rows = sequence ID, columns = features) of NA values

```{r}
na_rcs = list()
for (seqtype in seqtypes) {
  for (lambda in lambdas) {
    db_name = paste0(seqtype, lambda)
    print(db_name)
    
    isNaDB = is.na(featuresDB_orig[[db_name]])
    naInds = arrayInd(which(isNaDB), dim(isNaDB))
    na_rownames = featuresDB_orig[[db_name]][['rowname']][naInds[,1]]
    na_colnames = colnames(featuresDB_orig[[db_name]])[naInds[,2]]
    stopifnot(length(na_rownames) == length(na_colnames))
    
    if (length(na_rownames) > 0) {
      na_rcs[[db_name]] = data.frame(db = db_name, rownames = na_rownames, colnames = na_colnames)
    } else {
      na_rcs[[db_name]] = NULL
    }
  }
}
na_df = do.call(rbind, na_rcs)
na_df[['featureClass']] = sapply(strsplit(na_df$colnames, '_'), FUN = function(x) x[1])
na_df_byFeatureClass = na_df[!duplicated(na_df[c('db', 'rownames', 'type')]),]
```

```{r}
disorderSeqs = readRDS(paths[['disorderProteomeHuman']])[['disorderSeqs']]
d2p2ProteomeHuman = readr::read_tsv(paths[['d2p2ProteomeHuman']])
ensemblProteome = readr::read_tsv(paths[['ensemblProteome63']])
proteome = ensemblProteome %>%
  dplyr::rename(ensembl_peptide_id = id) %>%
  dplyr::select(ensembl_peptide_id, seq) %>%
  dplyr::right_join(d2p2ProteomeHuman, by = 'ensembl_peptide_id') %>%
  dplyr::select(ensembl_peptide_id, d2p2_id, seq) %>%
  dplyr::distinct(d2p2_id, .keep_all = TRUE) %>%
  dplyr::mutate(width = nchar(seq))

# map of full sequence IDs (d2p2_id) to width
widths_full = unlist(proteome[['width']])
names(widths_full) = proteome[['d2p2_id']]
fullIds = startsWith(na_df_byFeatureClass[['db']], 'full')
na_df_byFeatureClass[['width']][fullIds] = widths_full[na_df_byFeatureClass[['rownames']][fullIds]]

# map of disorder sequence IDs ([d2p2_id]_[start]) to width
widths_disorder = width(disorderSeqs)
names(widths_disorder) = names(disorderSeqs)
disorderIds = startsWith(na_df_byFeatureClass[['db']], 'disorder')
na_df_byFeatureClass[['width']][disorderIds] = widths_disorder[na_df_byFeatureClass[['rownames']][disorderIds]]

na_df_long = na_df_byFeatureClass %>% dplyr::filter(width > 2)
dim(na_df_long)
proteome %>% dplyr::filter(d2p2_id %in% na_df_long[['rownames']])
disorderSeqs[names(disorderSeqs) %in% na_df_long[['rownames']]]
```

Summary
- NA values only occur in full0, full25, disorder0, and disorder25
  - All NA values in full0 and disorder0 occur because the sequence length < 3
  - All NA values in full25 and disorder25 are caused by 4 proteins and 2 feature classes
    - Proteins: ENSP (D2P2_ID): seq
      - [ENSP00000441807](https://www.ensembl.org/Homo_sapiens/Transcript/Idhistory/Protein?t=ENSP00000441807) (D2P2 32711555): KEKKEERKKERKKEERKKERKKERKKEKERKKEKER
      - [ENSP00000446694](https://www.ensembl.org/Homo_sapiens/Transcript/Idhistory/Protein?t=ENSP00000446694) (D2P2 32708242): KRKKKERKKERKKKERKKERKKERKKEKERKKEKER
      - [ENSP00000449099](https://www.ensembl.org/Homo_sapiens/Transcript/Idhistory/Protein?t=ENSP00000449099) (D2P2 32710664): KRKKKERKKERRKKERKKERKKERKKEKERKKEKER
      - [ENSP00000449265](https://www.ensembl.org/Homo_sapiens/Transcript/Idhistory/Protein?t=ENSP00000449265) (D2P2 32706482): KEKKEERKKERKKKERKKERKKERKREKERKKEKER
    - Feature classes: Geary, Moran
    - These proteins are predicted to be fully disordered by VSL2b, so their disordered "subsequences" (32711555_1, 32708242_1, 32710664_1, 32706482_1), which are really just the full length sequences, also have NA-valued features.
    - All of these proteins are currently deprecated in Ensembl (Release 95). The last Ensembl release to include ENSP00000441807, ENSP00000449099, and ENSP00000449265 was Ensembl Release 75; ENSP00000446694 was last included in Ensembl Release 66.