---
title: "featurize sequences"
output: html_document
---

# Setup

```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "hold", fig.show = "hold")
library(readr)
library(dplyr)
library(reticulate)
# use_condaenv('rpy3_tmp')
Sys.setenv('RETICULATE_PYTHON'='/home/users/bentyeh/miniconda3/envs/rpy3_tmp/bin/python')
options(stringsAsFactors=FALSE, warn=1)

if (.Platform$OS.type == 'unix') {
    projectDir = normalizePath('~/projects/disprot')
} else {
    projectDir = normalizePath('~/Projects/disprot')
}
files = reticulate::import_from_path(module='files', path=file.path(projectDir, 'scripts'))
paths = files$getPaths()
source(file.path(projectDir, 'scripts', 'disorderedSeqs.R'))
source(file.path(projectDir, 'scripts', 'featurize.R'))
```

# Featurization

Script parameters
- `rebuildDisorderProteome`: rebuild disorder proteome (Ensembl proteome + D2P2 ID)
- `seqtypes`: 'full' (entire sequence); 'disorder' (only disordered regions)
- `lambdas`: only featurize sequences larger than these thresholds
- `features_sets`: features to use corresponding to `lambdas`
- `cores`: number of processes to use

```{r params}
rebuildDisorderProteome = FALSE
keepFeaturesInMemory = TRUE
seqtypes = c('full', 'disorder')
lambdas = c(0, 25, 50, 75, 100)
features_sets = list(setdiff(names(featureFuns_comp), 'TC'),
                     c(names(featureFuns_paac), names(featureFuns_cor)),
                     c(names(featureFuns_paac), names(featureFuns_cor)),
                     c(names(featureFuns_paac), names(featureFuns_cor)),
                     c(names(featureFuns_paac), names(featureFuns_cor)))
cores = 32
```

Build or read pre-existing proteome

```{r proteome, message = FALSE}
# Read in data
d2p2DisorderHuman = readr::read_tsv(paths[['d2p2DisorderHuman_vsl2b']])
d2p2ProteomeHuman = readr::read_tsv(paths[['d2p2ProteomeHuman']])
ensemblProteome = readr::read_tsv(paths[['ensemblProteome63']])

# proteome: modified Ensembl proteome
# - merge in D2P2 IDs: d2p2_id
# - rename peptide id column: ensembl_peptide_id
# - remove duplicated redundant sequences / D2P2 IDs (D2P2 IDs are non-redundant for sequences)
proteome = ensemblProteome %>%
  dplyr::rename(ensembl_peptide_id = id) %>%
  dplyr::select(ensembl_peptide_id, seq) %>%
  dplyr::right_join(d2p2ProteomeHuman, by = 'ensembl_peptide_id') %>%
  dplyr::select(ensembl_peptide_id, d2p2_id, seq) %>%
  dplyr::distinct(d2p2_id, .keep_all = TRUE)

if (rebuildDisorderProteome || !file.exists(paths[['disorderProteomeHuman']])) {
  print('Building disorder proteome...')
  disorderPred = d2p2DisorderHuman %>%
    dplyr::filter(d2p2_id %in% proteome[['d2p2_id']])
  tmp = extractDisorder(disorderPred = disorderPred, proteome = proteome,
                        disorder_idCol = 'd2p2_id', disorder_predCol = 'predictor_id',
                        proteome_seqCol = 'seq', proteome_idCol = 'd2p2_id',
                        uniqueNames = TRUE)
  saveRDS(tmp, file=paths[['disorderProteomeHuman']])
  disorderSeqs = tmp$disorderSeqs
} else {
  print(paste('Reading existing proteome:', paths[['disorderProteomeHuman']]))
  tmp = readRDS(paths[['disorderProteomeHuman']])
  disorderSeqs = tmp$disorderSeqs
}
```

Compute and save features

```{r featurize}
if (keepFeaturesInMemory) {
  featuresDB = list()
}

for (seqtype in seqtypes) {
  if (seqtype == 'full') {
    seqs = as.vector(proteome[['seq']])
    names(seqs) = proteome[['d2p2_id']]
  } else if (seqtype == 'disorder') {
    seqs = as.vector(disorderSeqs)
  }
  
  for (i in 1:length(lambdas)) {
    lambda = lambdas[[i]]
    features = features_sets[[i]]
    print(paste('Creating feature database:', paste('features', paste0(seqtype, lambda), 'human', sep='_')))
    
    mat = featurizeSeqs(seqs, lambda=lambda, features=features, cores=cores, format='matrix',
                        save=paths[[paste('features', paste0(seqtype, lambda), 'human', 'orig', sep='_')]])
    
    if (keepFeaturesInMemory) {
      featuresDB[[paste0(seqtype, lambda)]] = mat
    }
  }
}
```

