---
title: "visualizeData"
output: html_document
---

# Setup

```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "hold", fig.show = "hold")
library(readr)
library(dplyr)
library(ggplot2)
library(reticulate)
# use_condaenv("rpy3_tmp")
Sys.setenv("RETICULATE_PYTHON"="/home/users/bentyeh/miniconda3/envs/rpy3_tmp/bin/python")
options(stringsAsFactors=FALSE, warn=1)

if (.Platform$OS.type == "unix") {
    projectDir = normalizePath("~/projects/disprot")
} else {
    projectDir = normalizePath("~/Projects/disprot")
}
files <- import_from_path(module='files', path=file.path(projectDir, 'scripts'))
paths = files$getPaths()
source(file.path(projectDir, 'scripts', 'disorderedSeqs.R'))
source(file.path(projectDir, 'scripts', 'featurize.R'))
```

# Get Data

### Read in features database

`featuresDB`: mapping from character (`seqtype``lambda`) to matrix (rows = proteins, columns = features)

```{r featuresDB, message = FALSE}
featuresDB = list()
col_types = cols(.default = col_double(), rowname = col_character())
seqtypes = c('full', 'disorder')
lambdas = c(0, 25, 50, 75, 100)
for (seqtype in seqtypes) {
  for (lambda in lambdas) {
    print(paste0(seqtype, lambda))
    featuresDB[[paste0(seqtype, lambda)]] = as.matrix(
      readr::read_tsv(
        file = paths[[paste('features', paste0(seqtype, lambda), 'human', sep='_')]],
        col_types = col_types) %>%
      tibble::column_to_rownames()
    )
  }
}
```

### Gene lists

- Extract name of gene list (i.e., remove 'gl_' from the gene list name)
- Create directory in which to store plots: [gene list 1]_[gene list 2]

```{r gls}
gl_names = c('gl_TFs', 'gl_random')
gls = sapply(gl_names,
             function(gl_name) unlist(readr::read_tsv(paths[[gl_name]]) %>% dplyr::select(d2p2_id)))
gl_names_short = sub(pattern = 'gl_', replacement = '', x = gl_names)
plotsDir = file.path(paths[['dirResults']], paste0(gl_names_short, collapse='-'))
dir.create(plotsDir)
```

# Visualize Data

## Featurization

Script parameters
- `seqtypes`: 'full' (entire sequence); 'disorder' (only disordered regions)
- `lambdas`: sequence length thresholds
- `transformations`: 'PCA'; 't-SNE'
- `col_types`: column types for reading in features database

```{r params}
seqtypes = c('full', 'disorder')
lambdas = c(0, 25, 50, 75, 100)
transformations = c('PCA', 't-SNE')
```

Create visualization given complete features database (`featuresDB` list mapping from name (`seqtype``lambda`) --> features matrix)

```{r vis}
compareGeneListsAllFeatures(gls, featuresDB, seqtypes, lambdas, transformations, plotsDir)
```

